---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data

1. Load the data

```{r}
unzip("activity.zip")
activity <- read.csv("activity.csv")
```
2. Process/transform the data (if necessary) into a format suitable for your analysis

```{r}
library(dplyr)

steps_day <- activity %>% group_by(date) %>% summarise(steps_per_day = sum(steps))

steps_interval <- activity %>% group_by(interval) %>% summarise(avg_per_interval = mean(steps,na.rm=TRUE),med_per_interval = median(steps,na.rm=TRUE))


```



## What is mean total number of steps taken per day?

1. Make a histogram of the total number of steps taken each day

```{r}
hist(steps_day$steps_per_day)
```


2. Calculate and report the mean and median total number of steps taken per day
```{r}
mean(steps_day$steps_per_day,na.rm=TRUE)
median(steps_day$steps_per_day,na.rm=TRUE)
```


## What is the average daily activity pattern?

1. Make a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)

```{r}
with(steps_interval,plot(x=interval,y=avg_per_interval,type="l",xlab="5-minute interval",
     ylab="Average number of steps taken, averaged across all days",main="Time Series average number of steps taken over 5-minute intervals"))

max_avg_steps_interval <- steps_interval[which.max(steps_interval$avg_per_interval),]$interval
abline(v=max_avg_steps_interval,col="blue",lwd=1)
text(x=max_avg_steps_interval,y=max(steps_interval$avg_per_interval),labels=paste("max. number of steps at interval",max_avg_steps_interval),pos=4,col="blue")
points(x=max_avg_steps_interval,y=max(steps_interval$avg_per_interval),pch=16,lwd=3,col="blue")
```

2. Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?
```{r}
max_avg_steps_interval
```



## Imputing missing values



## Are there differences in activity patterns between weekdays and weekends?
